---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tickr

<!-- badges: start -->
![CRAN/METACRAN Version](https://img.shields.io/cran/v/tickr)
<!-- badges: end -->

This is a small package to provide consistent tick marks for plotting `ggplot2` figures. 
It provides breaks and labels for ggplot2 without requiring ggplot2 to be installed.

## Installation

You can install the development version of `tickr` from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("ben-williams/tickr")
```

## Example

```{r setup}
library(tickr)
library(ggplot2)
library(dplyr)
```

Some basic examples of how to use the `tickr` package.

```{r}
# Create a data.frame
df <- data.frame(year = 2000:2020,
                 spawning_biomass = rnorm(21,100, 0.25))

# explore tickr output
tickr(df, var=year)
tickr(df, var=year, by=2)
tickr(df, var=year, by=5, var_min = 1998, var_max = 2024, lab_start = 1998, lab_end = 2017)
tickr(df, var=year, by=5, var_min = 1998, var_max = 2024, lab_start = 1998, lab_end = 2017)

# explore plotting options
# default ggplot
ggplot(df, aes(year, spawning_biomass)) + 
  geom_line() +
  geom_point()

# a tick mark for every year
ggplot(df, aes(year, spawning_biomass)) + 
  geom_line() +
  geom_point() +
  scale_x_tickr(data=df, var=year)

# a tick mark for every year & label every 2nd year
ggplot(df, aes(year, spawning_biomass)) + 
  geom_line() +
  geom_point() +
  scale_x_tickr(data=df, var=year, by = 2)

# determine the years that are labeled
ggplot(df, aes(year, spawning_biomass)) + 
  geom_line() +
  geom_point() +
  scale_x_tickr(data=df, var=year, by = 5, 
                var_min = 1997, 
                var_max = 2024, 
                lab_start = 1997) 

# determine the years that are labeled
ggplot(df, aes(year, spawning_biomass)) + 
  geom_line() +
  geom_point() +
  scale_x_tickr(data=df, var=year, by = 2, 
                var_min = 2010, 
                lab_start = 2010)

# show future years
ggplot(df, aes(year, spawning_biomass)) + 
  geom_line() +
  geom_point() +
  scale_x_tickr(data=df, var=year, by = 5, 
                var_min = 2015, 
                var_max = 2026, 
                lab_start = 2015, 
                lab_end = 2026) +
  expand_limits(x = c(1998, 2026)) 

```

The same general approach is available for y axes as well.

```{r}
df <- expand.grid(age = 1:10,
                 year = 2000:2009) %>% 
  dplyr::mutate(value = rnorm(100, 1, 1))

df %>% 
  ggplot(aes(year, age, size=value)) + 
  geom_point() +
  scale_size_area() +
  scale_x_tickr(data=df, var=year, by=2, var_max = 2012, lab_end = 2012) +
  scale_y_tickr(data=df, var=age, by=2) +
  expand_limits(y=0, x=2012)


```

